Investigating Determinism of Physics Engines for usage in CAV and V\&V simulations

### 2021-10-20 rebuttal_letter/rebuttal.tex

The issue we found with taking the average deviation over many (100's or 1000's of) repeated runs is that a single `failure' can be hidden in an average. For example 1 error in 1000 would be an insignificant difference to an average, but a single simulation run that exceeds permissible variance could result in a false negative result or even fail to detect the presence of a serious fault with the system under test.

The verification process, therefore, needs to be aware of any deviation beyond the permissible tolerance, as even a single event may coincide with a bug or error in the system that needs to be found and corrected. Hence, it is imperative that we discover any simulation run that exceeds the permissible tolerance.

We could reframe this as thinking less about simulation performance, and more about error detection. It is important to know that any failure truly existed and if necessary can be repeated so that the bug can be tracked down and resolved. If the simulator is not deterministic, then by re-running the same test we may not be able to observe the same behaviour, i.e.\ we may not be able to recreate a previously seen error.

We have added (in italics) to the sentence on page 9 to clarify this point:  
\begin{quote} 
Initial testing [26] indicated an actor path deviation of $1\times10^{-13}$cm for 997 out of 1000 tests, with three tests reporting a deviation of over 10 $cm$. \textit{Due to the low probability of a simulation trace exceeding the permissible variance, using an average would `hide' these errors; hence this is the reason for using maximum variance and not an average}. 
\end{quote}

@@

\bigskip

\textit{Comment 3: "Moreover, I also concern about the results from several tests. As stated in Section V-B, 1 cm considered sufficient for urban scenario assertion checking. However, the average vehicle speeds are setting as 20–35 km/h (i.e. 5.56-9.72 m/s) [1]. Thus, I doubt the practical value of this permissible variance (too accurate for general AV testing) in real-world autonomous vehicle verification. The authors should provide more support materials for the setting used in the whole experiment"}.

Thank you we understand the point you are making. This seems to be less about the results quoted but more about what value of variance is appropriate for the given case study or application. The focus of this paper is not really on the absolute values that we chose for permissible variance, indeed we state this value should be set by the user in the methodology (see section VII.A.3), or the speed of the vehicles (which was based on the UK urban speed limit of 20 $mph$ or 30 $kmph$), but more about the method of detecting and controlling simulation variance in an end-user's system. The value of permissible variance chosen in the paper (1 $cm$) was considered an appropriate compromise between, say 1 $m$ (too coarse) and 1 $mm$ (too fine), to illustrate the violation of this limit given certain simulator and system conditions.

The point is that repeated runs should give the exact same output, regardless of the designated permissible variance or simulated vehicle speed. The reviewer suggests a paper that describes the speed of vehicles in an urban environment similar to those in our case study. We have included this reference together with a line to point interested readers who wish to set the speed of their simulated vehicles to an appropriate value.

### 2021-10-20 document/source/first_submit_body.tex
copy of first submitted body source

Design confidence is gained when the autonomous vehicle (AV) can be shown to comply with these rules e.g.,\ through assertion checking during simulation. There have been several fatalities with AVs, some of which could be attributed to insufficient verification and validation (V\&V), e.g.~\cite{FatalityExample}.

While on-road testing is an essential part of AV verification, it can be complemented by  simulation-based testing, which offers a means to explore the vast parameter space safely and efficiently~\cite{korosec2019waymo} whilst reducing the amount of costly road trials~\cite{kalra2016driving}.
%
%While on-road testing is an essential and complementary part of AV verification, simulation environments offer a means to explore the vast parameter space safely and efficiently~\cite{korosec2019waymo} whilst limiting the need for costly road trials~\cite{kalra2016driving}. 
In particular, simulations can be biased to increase the frequency at which otherwise rare events occur~\cite{Koopman2018}; this includes testing how the AV reacts to unexpected behaviour of the environment~\cite{RobustnessAutonomy}.

@@

In this paper, we investigate non-determinism and how it affects simulation results using the example of CARLA, an open-source autonomous driving simulation environment based on the Unreal Engine.
%
In our case study, scenarios between pedestrian and vehicle actors are investigated to determine the actor position variance in the simulation output for repeated simulation runs. 
%
We establish that the CARLA simulator is non-deterministic.  Actor path variance was found to be non-zero and, under certain conditions, a deviation from the mean was observed of up to $59$cm.
%
In such an urban environment, we consider a deviation of up to $1$cm to be permissible for AV verification.
%
However, in our experiments, CARLA only exhibits this permissible variance when system utilisation is restricted to 75\% or less and the simulation is terminated once a vehicle collision has been detected.

@@

This paper is structured as follows.
%
Section~\ref{s:prelim} defines terms used throughout the paper and identifies when determinism is needed. 
Section~\ref{s:background} briefly introduces how game engines work before investigating in Section~\ref{s:nondeterminisimSources} the potential sources of non-determinism in game engines.
%
An empirical case study of simulation variance for a number of scenarios involving pedestrians and vehicles is given in Section~\ref{s:case-study} including internal and external setting and system screening tests. 
%
The results from the case study are presented in Section~\ref{s:FinalResultsSection}. 
%
Section~\ref{s:methodology} presents the step-by-step method to assess the suitability of a simulation system for AV verification in general. 
%
We conclude in Section~\ref{s:conclusion} and give an outlook on future work.

@@

\subsubsection{Determinism}
Schumann et al.\ describe determinism as the property of causality given a temporal development of events such that any state is completely determined by prior states~\cite{Schumann2010}.\TODO{KIE: I can't find that quote in that paper, please point me to it.} However, in the context of simulation this should be expanded to include not just prior states but also the history of actions taken by all actors. Therefore, a deterministic simulation will always produce the same result given the same history of prior states and actions.

A simulation can be thought of as the process of generating or producing experimental data. In the case of a driving simulator, kinematics will describe future states of actors given the current conditions and actions taken, thereby generating new data. If a simulation is deterministic, Fig.~\ref{variance_description_b}, then there will be no variation in the generated output data, i.e. all future states are perfectly reproducible from prior states and actions. However, if a simulation is non-deterministic, Fig.~\ref{variance_description_a}, then there will be a variation in the generated output data. \\

\begin{figure}[!t]
    \centering

    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/Variance_predicition_tolerance_definition_diagram_a.pdf}
        \caption{Non-determinism}
        \label{variance_description_a}
    \end{subfigure}

    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/Variance_predicition_tolerance_definition_diagram_b.pdf}
        \caption{Determinism}
        \label{variance_description_b}
    \end{subfigure}

    \caption{Illustration of accuracy, variance, precision and tolerance in the context of non-determinism and determinism}
    \label{variance_description}
\end{figure}

@@

As an analogy, the simulator can be thought of as a manufacturing process that produces data. To determine the precision of this process, the output must be measured and analysed for differences when the process is repeated. Those differences describe the spread or variance in the process output. A hard limit on the variance can then be defined, Fig.~\ref{variance_description_a}, beyond which the output fails to meet the required tolerance, e.g.\ the output is rejected by quality control. Real manufacturing fails to achieve absolute precision. Hence, there is a need for tolerances to be specified to account for the variance in real-world manufacturing processes.

If a simulator is deterministic then it will produce results with absolute precision or zero variance, Fig.~\ref{variance_description_b}, and hence will be within tolerance by design. If the simulator is non-deterministic then there will be a measurable, non-zero variance in the output data.\\
%

@@

Considering the objectives for gaming and comparing them to those for AV development and testing, there are fundamental differences. Providing game players with a responsive real-time experience is often achieved at the cost of simulation accuracy and precision.
% 
The gamer neither needs a faithful representation of reality (i.e. gamer accepts low accuracy) nor requires repeated actions to result in the same outcome (i.e. gamer accepts low precision). In contrast, high accuracy and precision are necessary for AV development, testing and verification.

@@

A common misconception is to attribute non-determinism to the use of floating point number representation. This is an erroneous conflation with non-deterministic computational execution.  
%
Floating-point representation of real numbers is limited by the fixed bit width available in the hardware, resulting in the finite precision with which numbers can be represented in a computation. 
Thus, the results of arithmetic operations must fit into the given bit width, which is achieved through rounding to the nearest representable value. This gives rise to rounding errors~\cite{FloatingPointsBook,goldberg1991every}. %

As a consequence, floating-point arithmetic is not associative. Thus, for arithmetic expressions beyond two operands, such as $x+y+z$, executing $x+(y+z)$ rather than $(x+y)+z$, can produce different results~\cite{Kapre2007}. This is because rounding is being performed at different stages in the computation, depending on the respective execution order. 
%
% Translating to the usage of AV simulations in physics engines,
In the context of AV simulation, such rounding errors could result in accuracy issues of, for example, actor positions within the environment, leading to falsely satisfied or failed assertions.
%
Thus, one could conclude that floating-point arithmetic causes non-determinism, resulting in loss of repeatability. In fact, some authors suggest avoiding the use of floating-point representation entirely for assertion testing due to imprecision and non-deterministic behaviour~\cite{empirical-analysis-of-flaky-tests}.

@@

% However, as illustrated above on the example of non-associative addition, a change of execution order can lead to sequences of floating-point operations producing different results. Thus, any uncontrolled options to modify execution order, e.g.\ at the compiler level, within the runtime environment or at the hardware level, can cause simulation results to differ between runs, resulting in non-zero variance and loss of determinism.

% Beyond that, aggressive optimisations by the compiler can also introduce incorrectness in floating-point arithmetic~\cite{llvm-floating-point}. However, the same executable should still return the same output for identical input.

% Execution order changes can occur due to the use of a different compiler or as a consequence of parallelisation, e.g.\ within the runtime environment or at the hardware level. 
However, different compiler configurations, aggressive optimisations~\cite{llvm-floating-point}, and 
parallelisation within the runtime environment or at the hardware level can change the execution order, which can cause simulation results to differ between runs when floating-point operations are affected. 
%
Furthermore, floating-point calculation results may differ if the execution is performed on a GPU rather than a CPU which may have different register widths~\cite{Whitehead2011}.

@@

% ***************************************************
%  Scenario Description
% ***************************************************
\subsection{Context, Scene and Scenario}\label{TestsDescriptionAndTechnicalities}
\begin{figure}[!t]
    \centering
    \begin{subfigure}{.24\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/fig3a}
        \caption{}
        \label{Test_a}
    \end{subfigure}
    \begin{subfigure}{.24\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/det_fig3bb}
        \caption{}
        \label{Test_b}
    \end{subfigure}
    \begin{subfigure}{.24\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/det_fig3c}
        \caption{}
        \label{Test_c}
    \end{subfigure}
    \caption{Schematic of test scenarios for (a) Tests 1-2, (b) Tests 3-4, (c) Tests 5-6. Descriptions are given in Table~\ref{TableOfExperiments}.}\label{f:test_a_and_b}
\end{figure}

@@

For this case study a tolerance on actor position of $1$m would be insufficient when considering the spacial resolution required to distinguish between a collision and a near-miss event. A very small value, e.g.\ $1\times10^{-12}m$, may be overly-sensitive to minor computational perturbations and generate false positives. Therefore, for this case study a tolerance of $1$cm has been selected and thus any variance of less than $1$cm is permissible. To put this another way, we can accept a precision with a tolerance of $\leq$$\pm$1$cm$. This tolerance may need to be chosen for the specific verification case and the speed of the vehicles within the environment~\cite{jiang2014intercultural}.

@@

The tests used for the case study are listed in Table~\ref{TableOfExperiments}. They cover a range of interactions between actor types. The map size for the simulation test environment was 354$m$ $\times$ 170$m$. The range of movement of the actors within the environment was up to 70$m$ for vehicles and up to 25$m$ for pedestrians.
%
Tests 1 \& 2 involve two vehicles meeting at a junction where they either do not collide (Test 1) and where they do collide (Test 2), thereby triggering an \textit{event hit} callback in the game engine. In both cases the trajectories of the vehicles are hard-coded to follow a set of waypoints spaced at $0.1$m intervals using a PID controller. 
%
In Test 3 a mixture of different actor types is introduced where two vehicles drive without collision and a pedestrian walks across the road at a crossing point.

@@

Instead of using variance, herein the results are given in terms of the deviation, $\sigma_a(t)$, which indicates the dispersion of the actor path relative to the mean and is helpfully in the same units as actor position, i.e.\ metres ($m$), for ease of interpretation. The maximum variance ($\Psi$) over the entire set of $n$ repeated simulations, i.e.\ the overall observed worst case, is defined as the largest variance of any actor at any time in any of the simulation runs, as given in Equation~\ref{eq:max_sigma}. 
\begin{equation} \label{eq:max_sigma}
\Psi = \max_{a,t}\sigma_a^{2}(t)
\end{equation}
The maximum deviation is the absolute value of the square root of the maximum variance and herein referred to as ${\max\sigma}$ for brevity.

@@

\section{Results and Discussion}\label{s:FinalResultsSection}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.48\textwidth]{../other/figures/actor_path_plot.jpg}
    \caption{Actor path plot in XY plane for Test 4 with 95\% resource utilisation.}
    \label{actorPathPlot}
\end{figure}

@@

Initial testing~\cite{TSLUnrealEngineTesting} indicated an actor path deviation of $1\times10^{-13}$cm for 997 out of 1000 tests, with three tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10$cm. While executing 100 repeats may seem sufficient, this sample size may fail to observe events that occur with low probability, giving false confidence in the results. Therefore each test was repeated 1000 times to provide a sufficient sample size. 
%
Due to the low probability of a simulation trace exceeding the permissible variance, using an average would `hide' these errors; hence this is the reason for using maximum variance and not an average.
%
A detailed guide for reproducing the experiments along with the scripts used are provided on github\footnote{\url{https://github.com/TSL-UOB/CAV-Determinism/tree/master/CARLA_Tests_setup_guide}}. To eliminate some of the potential sources of non-determinism outlined in Section~\ref{s:nondeterminisimSources} a series of screening tests and analyses were performed on our system. These were:

@@

% ***************************************************
%  Results and Discussion
% ***************************************************

@@

A summary of the main results are shown in Table~\ref{TableOfExperiments}.  In the column $\max\sigma$ (unrestricted) the value reported is the maximum deviation across all resource utilisation levels, i.e.\ the worst case for a given scenario. From these results it is clear that scenarios with only pedestrian actors (Tests 5 \& 6) display results within tolerance over all resource utilisation levels with or without a collision where $\max\sigma$ is $5.6\times10^{-13}$ or $0.56\si{\pico\metre}$. However, all other scenarios involving vehicles or a mixture of actor types do not meet the required tolerance, with some deviation in actor path as large as $59$cm. 
%
A plot of actor position in the X-Y plane (plan view, units $m$) is shown in Fig.\ref{actorPathPlot}, where the inset clearly shows the divergence of the path of vehicle 2 post-collision with the pedestrian.
%
Clearly, such a large deviation cannot be acceptable for simulation to be considered a credible verification tool.

>>>>

Design confidence is gained when the autonomous vehicle (AV) can be shown to comply with these rules e.g.,\ through assertion checking during simulation. There have been several fatalities with AVs, some of which could be attributed to insufficient verification and validation (V\&V), e.g.~\cite{FatalityExample}. Simulation environments offer a means to explore the vast parameter space safely and efficiently~\cite{korosec2019waymo} without the need for millions of miles of costly on-road testing~\cite{kalra2016driving}. In particular, simulations can be biased to increase the frequency at which otherwise rare events occur~\cite{Koopman2018}; this includes testing how the AV reacts to unexpected behaviour of the environment~\cite{RobustnessAutonomy}.

@@

In this paper, we investigate non-determinism and how it affects simulation results using the example of CARLA, an open-source autonomous driving simulation environment based on the Unreal game engine.
%
In our case study, scenarios between pedestrian and vehicle actors are investigated to determine the actor position variance in the simulation output for repeated simulation runs. We find that the CARLA simulator is non-deterministic. Actor path variance was found to be non-zero and, under certain conditions, the deviation from the mean was observed up to $59$cm. We would consider a deviation of $1$cm to be permissible for autonomous vehicle verification tasks in an urban environment. However, we found that CARLA exhibits a permissible variance when system utilisation is restricted to 75\% or less and the simulation is terminated once a vehicle collision has been detected. 
%

@@

This paper is structured as follows.
%
Section~\ref{s:prelim} defines terms used throughout the paper and identifies when determinism is needed. 
Section~\ref{s:background} briefly introduces how game engines work before investigating in Section~\ref{s:nondeterminisimSources} the potential sources of non-determinism in game engines.
%
An empirical case study of simulation variance for a number of scenarios involving pedestrian and vehicles is given in Section~\ref{s:case-study} including internal and external setting and system screening tests. 
%
The results from the case study are presented in Section~\ref{s:FinalResultsSection}. 
%
Section~\ref{s:methodology} presents the step-by-step method to assess the suitability of a simulation system for AV verification in general. 
%
We conclude in Section~\ref{s:conclusion} and give an outlook on future work.

@@

\subsubsection{Determinism}
Schumann describes determinism as the property of causality given a temporal development of events such that any state is completely determined by prior states~\cite{Schumann2010}. However, in the context of simulation this should be expanded to include not just prior states but also the history of actions taken by all actors. Therefore, a deterministic simulation will always produce the same result given the same history of prior states and actions.

A simulation can be thought of as the process of generating or producing experimental data. In the case of a driving simulator, kinematics will describe future states of actors given the current conditions and actions taken, thereby generating new data. If a simulation is deterministic, Fig.~\ref{variance_description}~(b), then there will be no variation in the generated output data, i.e. all future states are perfectly reproducible from prior states and actions. However, if a simulation is non-deterministic, Fig.~\ref{variance_description}~(a), then there will be a variation in the generated output data. \\

\begin{figure}[!t]
    \centering
    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/Variance_predicition_tolerance_definition_diagram_a.pdf}
        \caption{Non-deterministic}
        \label{variance_description_a}
    \end{subfigure}

    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/Variance_predicition_tolerance_definition_diagram_b.pdf}
        \caption{Deterministic}
        \label{variance_description_b}
    \end{subfigure}
    \caption{Demonstration of variance, precision, tolerance and determinism}
    \label{variance_description}
\end{figure}

@@

As an analogy, the simulator can be thought of as a manufacturing process that produces data. To determine the precision of this process, the output must be measured and analysed for differences when the process is repeated. Those differences describe the spread or variance in the process output. A hard limit on the variance can then be defined, Fig.~\ref{variance_description}~(a), beyond which the output fails to meet the required tolerance, e.g.\ the output is rejected by quality control. Real manufacturing fails to achieve absolute precision. Hence, there is a need for tolerances to be specified to account for the variance in real-world manufacturing processes.

If a simulator is deterministic then it will produce results with absolute precision or zero variance, Fig.~\ref{variance_description}~(b), and hence will be within tolerance by design. If the simulator is non-deterministic then there will be a measurable, non-zero variance in the output data.\\
%

@@

Considering the objectives for gaming and comparing them to these for AV development and testing, there are fundamental differences. Providing game players with a responsive real-time experience is often achieved at the cost of simulation accuracy and precision.
% 
The gamer neither needs a faithful representation of reality (i.e. gamer accepts low accuracy) nor require repeated actions to result in the same outcome (i.e. gamer accepts low precision). In contrast, high accuracy and precision are necessary for AV development, testing and verification.

@@

Floating-point representation of real numbers is limited by the fixed bit width available in the hardware, resulting in the finite precision with which numbers can be represented in a computation. 
Thus, the results of arithmetic operations must fit into the given bit width, which is achieved through rounding to the nearest representable value. This gives rise to rounding errors~\cite{FloatingPointsBook,goldberg1991every}. %

As a consequence, floating-point arithmetic is not associative. Thus, for arithmetic expressions beyond two operands, such as $x+y+z$, executing $x+(y+z)$ rather than $(x+y)+z$, can produce different results~\cite{Kapre2007}. This is because rounding is being performed at different stages in the computation, depending on the respective execution order. 
%
Execution order changes can occur due to the use of a different compiler or as a consequence of parallelisation, e.g.\ within the runtime environment or at the hardware level. 
%
Furthermore, floating-point calculation results may differ if the execution is performed on a GPU rather than a CPU which may have different register widths~\cite{Whitehead2011}.

% Translating to the usage of AV simulations in physics engines,
In the context of AV simulation, such rounding errors could result in accuracy issues of, for example, actor positions within the environment, leading to falsely satisfied or failed assertions.
%
Thus, one could conclude that floating-point arithmetic causes non-determinism, resulting in loss of repeatability. In fact, some authors suggest avoiding the use of floating-point representation entirely for assertion testing due to imprecision and non-deterministic behaviour~\cite{empirical-analysis-of-flaky-tests}.

@@

However, as illustrated above on the example of non-associative addition, a change of execution order can lead to sequences of floating-point operations producing different results. Thus, any uncontrolled options to modify execution order, e.g.\ at the compiler level, within the runtime environment or at the hardware level, can cause simulation results to differ between runs, resulting in non-zero variance and loss of determinism.

Beyond that, aggressive optimisations by the compiler can also introduce incorrectness in floating-point arithmetic~\cite{llvm-floating-point}. However, the same executable should still return the same output for identical input.

@@

% ***************************************************
%  Scenario Description
% ***************************************************
\subsection{Context, Scene and Scenario}\label{TestsDescriptionAndTechnicalities}
\begin{figure}[!t]
    \centering
    \begin{subfigure}{.24\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/TestCasesDiagramV2_a.pdf}
        \caption{}
        \label{Test_a}
    \end{subfigure}
    \begin{subfigure}{.24\textwidth}
        \includegraphics[width=1\textwidth]{../other/figures/TestCasesDiagramV2_b.pdf}
        \caption{}
        \label{Test_b}
    \end{subfigure}
    \caption{Schematic of test scenarios for (a) Tests 1-4, (b) Tests 5-6. Descriptions are given in Table~\ref{TableOfExperiments}.}\label{f:test_a_and_b}
\end{figure}

@@

For this case study a tolerance on actor position of $1$m would be insufficient when considering the spacial resolution required to distinguish between a collision and a near-miss event. A very small value, e.g.\ $1\times10^{-12}m$, may be overly-sensitive to minor computational perturbations and generate false positives. Therefore, for this case study and, more broadly, for our general verification requirements, a tolerance of $1$cm has been selected. Thus any variance of less than $1$cm is permissible. To put this another way, we can accept a precision with a tolerance of $\leq$$\pm$1$cm$.

@@

The tests used for the case study are listed in Table~\ref{TableOfExperiments}. They cover a range of interactions between actor types. 
%
Tests 1 \& 2 involve two vehicles meeting at a junction where they either do not collide (Test 1) and where they do collide (Test 2), thereby triggering an \textit{event hit} callback in the game engine. In both cases the trajectories of the vehicles are hard-coded to follow a set of waypoints spaced at $0.1$m intervals using a PID controller. 
%
In Test 3 a mixture of different actor types is introduced where two vehicles drive without collision and a pedestrian walks across the road at a crossing point.

@@

Instead of using variance, herein the results are given in terms of the deviation, $\sigma_a(t)$, which indicates the dispersion of the actor path relative to the mean and is helpfully in the same units as actor position, i.e.\ metres ($m$), for ease of interpretation. The maximum variance over the entire set of $n$ repeated simulations, i.e.\ the overall observed worst case, is defined as the largest variance of any actor at any time in any of the simulation runs, as given in Equation~\ref{eq:max_sigma}. 
\begin{equation} \label{eq:max_sigma}
\max_{a,t}\sigma_a^{2}(t)
\end{equation}
The maximum deviation is the absolute value of the square root of the maximum variance and herein referred to as ${\max\sigma}$ for brevity.

@@

@@

Initial testing~\cite{TSLUnrealEngineTesting} indicated an actor path deviation of $1\times10^{-13}$cm for 997 out of 1000 tests, with three tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10$cm. While executing 100 repeats may seem sufficient, this sample size may fail to observe events that occur with low probability, giving false confidence in the results. Therefore each test was repeated 1000 times to provide a sufficient sample size. A detailed guide for reproducing the experiments along with the scripts used are provided on github\footnote{\url{https://github.com/TSL-UOB/CAV-Determinism/tree/master/CARLA_Tests_setup_guide}}. To eliminate some of the potential sources of non-determinism outlined in Section~\ref{s:nondeterminisimSources} a series of screening tests and analyses were performed on our system. These were:

@@

% ***************************************************
%  Results and Discussion
% ***************************************************
\section{Results and Discussion}\label{s:FinalResultsSection}

@@

A summary of the main results are shown in Table~\ref{TableOfExperiments}.  In the column $\max\sigma$ (unrestricted) the value reported is the maximum deviation across all resource utilisation levels, i.e.\ the worst case for a given scenario. From these results it is clear that scenarios with only pedestrian actors (Tests 5 \& 6) display results within tolerance over all resource utilisation levels with or without a collision where $\max\sigma$ is $5.6\times10^{-13}$ or $0.56\si{\pico\metre}$. However, all other scenarios involving vehicles or a mixture of actor types do not meet the required tolerance, with some deviation in actor path as large as $59$cm. Clearly, such a large deviation cannot be acceptable for simulation to be considered a credible verification tool.
